Unshuffling Permutations

### 2015-09-14 article.tex
Some typos and improvements

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
\usepackage{enumitem}
\usepackage{shuffle}

@@

% \renewcommand{\floatpagefraction}{.9}
% \renewcommand{\textfraction}{.1}

@@

We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations are specified by
words without repetition of length $n$ on the alphabet $[n]$, so that
all above definitions about words remain valid on permutations. If
$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
integers without multiple occurrences of a same letter, $\STD(u)$ is the
{\em standarized} of $u$, that is the unique permutation of the same
size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
is the set $S$ of all permutations. Two words $u$ and $v$ having the
same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
of $\sigma$, $\pi$ {\em avoids} $\sigma$.

@@

Let use denote by $\QQ[S]$ the linear span of all permutations. As
announced, we shall define on this space a kind of shuffle product and
for this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
For any permutation $\pi$, we set
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P_1 \uplus P_2 = [|\pi|]}
    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
three ways to extract two disjoint subwords of the permutation $1432$
so that the letters of these two subwords have respectively the same
relative order as the letters of the permutations $1$ and $132$.
\medskip

As any coproduct, $\Delta$ leads to the definition of a product obtained
by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
for any permutation $\pi$, we have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$ denoted by
$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
permutations
$\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

Intuitively, this product shuffles the values and the positions of the
letters of the permutations. There are other shuffle products on
permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
{\em convolution product}. Our shuffle product is different from these
last two.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
zero. In a more combinatorial way, this is equivalent to say that there
are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
is order isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is order
isomorphic to $\nu$.
\medskip

Besides, a permutation $\pi$ is a {\em square} if there is a
permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
if in the expansion of $\Delta(\pi)$, there is a tensor
$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
combinatorial way, this is equivalent to say that there are two sets
$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
$\pi_{|P_2}$ are order isomorphic. Computer experiments give us the
first numbers of square permutations with respects to their size, which
are, from size $0$ to $10$,
\begin{equation}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation}
For instance, the square permutations of sizes $0$ to $4$ are
\begin{multline}
    \epsilon, 12, 21,
    1234, 1243, 1423, 1324, 1342, 4132, 3124, 3142, \\ 3412, 4312,
    2134, 2143, 2413, 4213, 2314, 2431, 4231, 3241, 3421, 4321.
\end{multline}
\medskip

@@

        Finally, observe that any permutation $\pi$ avoiding $213$ and
        $231$ necessarily starts by the smallest possible letter or the
        greatest possible letter. This property is then true for the
        suffix of $\pi$ obtained by deleting its first letter,
        and so on for all of its suffixes. Thus, by
        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
        role of a smallest (resp. greatest) letter, one obtains a binary
        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_binary_to_square_permutation}
        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
        permutation.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
        Since $u$ is a square binary word, there is a binary word $v$
        such that $u \in v \shuffle v$. Then, there are two disjoint
        sets $P$ and $Q$ of positions of letters of $u$ such that
        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
        same standarized $\rho$. Hence, and by definition of the shuffle
        product of permutations, $\BINTOPERM(u)$ appears in
        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
        permutation.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_permutation_to_square_binary}
        If $\pi$ is a square permutation avoiding $213$ and $231$,
        $\PERMTOBIN(\pi)$ is a square binary word.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
        well-defined binary word. Since $\pi$ is a square permutation,
        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are order
        isomorphic. This implies, by the definitions of $\BINTOPERM$ and
        $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$ is a
        square binary word.
        \qed
    \end{proof}
\end{proof}
\medskip

@@

\begin{proposition} \label{prop:shuffle_associative_commutative}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
    To prove the associativity of $\SHUFFLE$, it is convenient to show
    that its dual coproduct $\Delta$ is coassociative, that is
    \begin{equation}
        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
    \end{equation}
    where $I$ denote the identity map. This strategy relies on the fact
    that a product is associative if and only if its dual coproduct is
    coassociative. For any permutation $\pi$, we have
    \begin{equation} \begin{split}
    \label{equ:shuffle_associative_commutative}
        (\Delta \otimes I) \Delta(\pi) & =
        (\Delta \otimes I)
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]} \;
        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
        \otimes
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
        \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes
        \STD\left(\pi_{|P_2}\right) \otimes
        \STD\left(\pi_{|P_3}\right).
    \end{split} \end{equation}
    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
    is equal to the last member
    of~\eqref{equ:shuffle_associative_commutative}, whence the
    associativity of $\SHUFFLE$.
    \smallskip

@@

    If $P$ is a set of indexes of letters of $\pi$, we denote by
    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
    the operation $\widetilde{\,}$ define a bijection on the set of the
    subsets of $[|\pi|]$, and since the standardization operation
    commutes with the mirror operation on words, we have
    \begin{equation} \begin{split}
        \Delta(\phi_1(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_1(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|P_1}\right)
        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \widetilde{\STD\left(\pi_{|P_1}\right)}
        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
    $\phi_1$ is a morphism of associative algebras.
    \smallskip

    Next, since by definition of the complementation operation on
    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_2(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_2(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\bar \pi_{|P_1}\right)
        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
    $\phi_2$ is a morphism of associative algebras.
    \smallskip

>>>>

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
\usepackage{enumitem}
\usepackage{shuffle}
\usepackage{multirow}

@@

% \renewcommand{\floatpagefraction}{.9}
% \renewcommand{\textfraction}{.1}
\renewcommand{\arraystretch}{1.2}

@@

For any nonnegative integer $n$, $[n]$ is the set $\{1, \dots, n\}$.
We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations of a size $n$
are specified by words of length $n$ on the alphabet $[n]$ and without
multiple occurrence of a letter, so that all above definitions about
words remain valid on permutations. If $\pi$ is a permutation, we denote
by $\bar \pi$ (resp.~$\pi^{-1}$) the {\em complementary} (resp.
{\em inverse}) of $\pi$. If $u$ is a word of integers without multiple
occurrences of a same letter, $\STD(u)$ is the {\em standarized} of $u$,
that is the unique permutation of the same size as $u$ such that for all
$i, j \in [|u|]$, $u_i < u_j$ if and only if $\STD(u)_i < \STD(u)_j$. In
particular, the image of the map $\STD$ is the set $S$ of all
permutations. Two words $u$ and $v$ having the same standarized are
{\em order-isomorphic}. If $\sigma$ is a permutation, there is an
{\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if there is a
set $P$ of indexes of letters of $\pi$ such that $\sigma$ and $\pi_{|P}$
are order-isomorphic. When $\pi$ does not admit any occurrence of
$\sigma$, $\pi$ {\em avoids} $\sigma$.

@@

Let use denote by $\QQ[S]$ the linear span of all permutations. We
define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
any permutation $\pi$, we set
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P_1 \sqcup P_2 = [|\pi|]}
    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
three ways to extract from the permutation $1432$ two disjoint subwords
respectively order-isomorphic to the permutations $1$ and $132$.
\medskip

As announced, let us now use $\Delta$ to define a shuffle product on
permutations. As any coproduct, $\Delta$ leads to the definition of a
product obtained by duality in the following way.
From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$, denoted by
$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
any permutations $\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

Intuitively, this product shuffles the values and the positions of the
letters of the permutations. As first and easy properties of $\SHUFFLE$,
one can observe that the empty permutation $\epsilon$ is a unit for
$\SHUFFLE$ and that this product is graded by the sizes of the
permutations ({\em i.e.}, the product of a permutation of size $n$ with
a permutation of size $m$ produces a sum of permutations of size $n + m$.
Notice that there are other shuffle products on
permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
{\em convolution product}. Our shuffle product is different from these
last two.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
zero. In a more combinatorial way, this is equivalent to say that there
are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
is order-isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is
order-isomorphic to $\nu$.
\medskip

Besides, a permutation $\pi$ is a {\em square} if there is a
permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
if in the expansion of $\Delta(\pi)$, there is a tensor
$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
combinatorial way, this is equivalent to say that there are two sets
$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
$\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
first numbers of square permutations with respects to their size, which
are, from size $0$ to $10$,
\begin{equation}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation}
For instance, the square permutations of sizes $0$ to $4$ are
\smallskip

\begin{tabular}{c|c|c}
    Size $0$ \, & Size $2$ & Size $4$ \\ \hline
    \multirow{2}{*}{\, $\epsilon$} &
    \multirow{2}{*}{\, $12$, $21$ \,} &
    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
    $3412$, $4312$, \\
    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
    $3421$, $4321$
\end{tabular}
\medskip

@@

        Finally, observe that any permutation $\pi$ avoiding $213$ and
        $231$ necessarily starts by the smallest possible letter or the
        greatest possible letter. This property is then true for the
        suffix of $\pi$ obtained by deleting its first letter,
        and so on for all of its suffixes. Thus, by
        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
        role of a smallest (resp. greatest) letter, one obtains a binary
        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_binary_to_square_permutation}
        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
        permutation.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
        Since $u$ is a square binary word, there is a binary word $v$
        such that $u \in v \shuffle v$. Then, there are two disjoint
        sets $P$ and $Q$ of positions of letters of $u$ such that
        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
        same standarized $\rho$. Hence, and by definition of the shuffle
        product of permutations, $\BINTOPERM(u)$ appears in
        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
        permutation.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_permutation_to_square_binary}
        If $\pi$ is a square permutation avoiding $213$ and $231$,
        $\PERMTOBIN(\pi)$ is a square binary word.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
        well-defined binary word. Since $\pi$ is a square permutation,
        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
        order-isomorphic. This implies, by the definitions of $\BINTOPERM$
        and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
        is a square binary word.
        \qed
    \end{proof}
\end{proof}
\medskip

@@

\begin{proposition} \label{prop:shuffle_associative_commutative}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
    To prove the associativity of $\SHUFFLE$, it is convenient to show
    that its dual coproduct $\Delta$ is coassociative, that is
    \begin{equation}
        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
    \end{equation}
    where $I$ denote the identity map. This strategy relies on the fact
    that a product is associative if and only if its dual coproduct is
    coassociative. For any permutation $\pi$, we have
    \begin{equation} \begin{split}
    \label{equ:shuffle_associative_commutative}
        (\Delta \otimes I) \Delta(\pi) & =
        (\Delta \otimes I)
        \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & =
        \sum_{P_1 \sqcup P_2 = [|\pi|]} \;
        \sum_{Q_1 \sqcup Q_2 = [|P_1|]}
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
        \otimes
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
        \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \sqcup P_2 \sqcup P_3 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes
        \STD\left(\pi_{|P_2}\right) \otimes
        \STD\left(\pi_{|P_3}\right).
    \end{split} \end{equation}
    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
    is equal to the last member
    of~\eqref{equ:shuffle_associative_commutative}, whence the
    associativity of $\SHUFFLE$.
    \smallskip

@@

    If $P$ is a set of indexes of letters of $\pi$, we denote by
    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
    the operation $\widetilde{\,}$ defines a bijection on the set of the
    subsets of $[|\pi|]$, and since the standardization operation
    commutes with the mirror operation on words without multiple
    occurrence of a letter, we have
    \begin{equation} \begin{split}
        \Delta(\phi_1(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_1(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|P_1}\right)
        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \widetilde{\STD\left(\pi_{|P_1}\right)}
        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
    $\phi_1$ is a morphism of associative algebras.
    \smallskip

    Next, since by definition of the complementation operation on
    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_2(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_2(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\bar \pi_{|P_1}\right)
        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
    $\phi_2$ is a morphism of associative algebras.
    \smallskip

    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_3(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_3(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|P_1}\right)
        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right)^{-1}
        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
    $\phi_3$ is a morphism of associative algebras.
    \qed
\end{proof}
\medskip
